---
title: "Data Science Capstone - Milestone Report"
author: "Daniel Contrera"
date: "12/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Getting and Loading the data

The following zip file contains the training data:
https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip

```{bash, eval = FALSE}
unzip("Coursera-SwiftKey.zip")
```

There will be the following folders and files:
```{bash, eval = FALSE}
./Coursera-SwiftKey:
final
./Coursera-SwiftKey/final:
de_DE  en_US  fi_FI  ru_RU
./Coursera-SwiftKey/final/de_DE:
de_DE.blogs.txt  de_DE.news.txt  de_DE.twitter.txt
./Coursera-SwiftKey/final/en_US:
en_US.blogs.txt  en_US.news.txt  en_US.twitter.txt
./Coursera-SwiftKey/final/fi_FI:
fi_FI.blogs.txt  fi_FI.news.txt  fi_FI.twitter.txt
./Coursera-SwiftKey/final/ru_RU:
ru_RU.blogs.txt  ru_RU.news.txt  ru_RU.twitter.txt
```

Each folder has three files blogs, new, and twitter. We are going to start using the ones from `en_US` folder which correspond to English (US).

To start the exploration analysis we can open the files in a text editor, command line or load them in RStudio.
We loaded them in RStudio because we are getting familiar with the text management in R.

To perform a first exploration in the command line you can run the following:

```{bash, eval = FALSE}
# How many bytes
Coursera-Swiftkey/final/en_US$ ls -alh 
```

ReadLines in R:
```{r, cache = TRUE, eval = FALSE, warning = FALSE, message = FALSE}
length(readLines("Coursera-Swiftkey/final/en_US/en_US.twitter.txt")) # 2360148
length(readLines("Coursera-Swiftkey/final/en_US/en_US.blogs.txt")) # 899288
length(readLines("Coursera-Swiftkey/final/en_US/en_US.news.txt")) # 1010242
```

What is the length of the longest line seen in any of the three en_US data sets? 
```{bash, cache = TRUE, eval = TRUE}
wc -L Coursera-Swiftkey/final/en_US/*.txt
```

<!-- ```{bash} -->
<!-- wc −l Coursera-Swiftkey/final/en_US/en_US.twitter.txt -->
<!-- ``` -->

We will use the following function called `readText()` to read the files in R:

```{r}
readText <- function (
    filename, 
    nchunk = 1, 
    MAXLINES = 3e9, 
    words = c(), 
    search = ""
) {
    nlines <- 0
    lines <- "lines"
    longest <- 0
    w1 <- 0
    w2 <- 0 
    found <- c()
    con <- file(filename, "r") 
    while (length(lines) > 0 & nlines < MAXLINES) {
        lines <- readLines(con, nchunk, ok = TRUE)
        # print(lines)
        longest <- max(longest, nchar(lines))
        nlines <- nlines + length(lines)
        if (length(words) == 2) {
            w1 <- w1 + length(grep(words[1], lines))
            w2 <- w2 + length(grep(words[2], lines))
        }
        if (nchar(search) > 0 & length(grep(search, lines))) {
            found <- c(found, lines[grep(search, lines)])
        }
    }
    close(con)
    text <- c()
    text$nlines <- nlines
    text$longest <- longest
    text$w1 <- w1
    text$w2 <- w2
    text$found <- found
    text
}
```
This function reads a file (in chunks and lines can be limited) and returns the 
number of lines, the longest lines and how many words match a tuple input.

```{r, cache = TRUE, eval = FALSE, warning=FALSE, message=FALSE}
filenames <- c(
    "Coursera-SwiftKey/final/en_US/en_US.twitter.txt",
    "Coursera-SwiftKey/final/en_US/en_US.blogs.txt",
    "Coursera-SwiftKey/final/en_US/en_US.news.txt"
)
for (f in filenames) {
    text <- readText(f, nchunk = 1000)
    print(f)
    print(text)
}
```

## Sampling the training data

We are going to sample the training data because it's too large and we can infer many things from a smaller subset. We are going to use 10000 lines from each file and save them to another file.

The following function performs the sampling:
```{r, cache = TRUE, eval=FALSE, message = FALSE, warning=FALSE}
sampleText <- function (f, MAXLINES = 100, nchunk = 1, p = 0.5, pTrain = 1) {
    f_split <- strsplit(f, "/")[[1]]
    con <- file(f, "r")
    # f_out <- paste(f, "samp", MAXLINES, nchunk, p, sep = '.')
    f_out <- paste("Coursera-SwiftKey/train/", f_split[length(f_split)], sep = "")
    f_out_test <- paste("Coursera-SwiftKey/test/", f_split[length(f_split)], sep = "")
    con_out <- file(f_out, "w")
    con_out_test <- file(f_out_test, "w")
    nlines <- 0
    lines <- "lines"
    while (length(lines) > 0 & nlines < MAXLINES) {
        lines <- readLines(con, nchunk, ok = TRUE)
        # print(lines)
        if (rbinom(1, 1, p)) {
            if (pTrain < 1 & rbinom(1, 1, 1 - pTrain)) {
                    writeLines(lines, con = con_out_test)
            } else {
                writeLines(lines, con = con_out)
                nlines <- nlines + length(lines)
            }
        }
    }
    close(con)
    close(con_out)
    close(con_out_test)
}
filenames <- c(
    "Coursera-SwiftKey/final/en_US/en_US.twitter.txt",
    "Coursera-SwiftKey/final/en_US/en_US.blogs.txt",
    "Coursera-SwiftKey/final/en_US/en_US.news.txt"
)
max_lines <- 10000
nchunk <- 1
p <- 1/(500000/max_lines)
pTrain <- 0.8
for (f in filenames) {
    sampleText(f, max_lines, nchunk, p, pTrain)
}
```
So, we then have to read from the subset files.

## Exploratory Data Analysis

In the following chunk we read each subset
```{r, cache = TRUE, eval = TRUE, message = FALSE, warning=FALSE}
filenames <- c(
    # "Coursera-SwiftKey/final/en_US/en_US.twitter.txt",
    # "Coursera-SwiftKey/final/en_US/en_US.blogs.txt",
    # "Coursera-SwiftKey/final/en_US/en_US.news.txt"
    # "Coursera-SwiftKey/final/en_US/en_US.twitter.txt.samp.10000.1.0.02",
    # "Coursera-SwiftKey/final/en_US/en_US.blogs.txt.samp.10000.1.0.02",
    # "Coursera-SwiftKey/final/en_US/en_US.news.txt.samp.10000.1.0.02"
    "Coursera-SwiftKey/train/en_US.twitter.txt",
    "Coursera-SwiftKey/train/en_US.blogs.txt",
    "Coursera-SwiftKey/train/en_US.news.txt"
)
lines_file <- c()
for (f in filenames) {
    con <- file(f, "r")
    max_lines <- 10000
    lines_file <- c(lines_file, readLines(con, max_lines))
    close(con)
}
max_lines <- length(lines_file)
```


Then:

- keep the letters, `'` and spaces

- replace double spaces with a single one

- delete spaces at the beginning or end

- lowercase

```{r, cache = TRUE, eval = TRUE}
inds <- 1:max_lines
lines <- lines_file[inds]
lines[inds] <- gsub("((?![a-zA-Z' ]).)*", "", lines[inds], perl = TRUE)
lines[inds] <- gsub("  +", " ", lines[inds])
lines[inds] <- gsub("^ | $", "", lines[inds])
lines[inds] <- tolower(lines[inds])
```

Calculate the frequencies
```{r, cache=TRUE, eval = FALSE}
# words
words <- Reduce(paste, lines[inds])
words <- strsplit(words, " ")[[1]]
words <- words[sort.list(words)]
# frequency table
t <- table(words)
t <- t[sort.list(as.vector(t), decreasing = TRUE)]
t <- as.data.frame(t)
t$perc <- t$Freq/(sum(t$Freq))
t$perc_cum <- t$perc
for (i in 2:length(t$perc_cum)) {
    t$perc_cum[i] <- t$perc_cum[i - 1] + t$perc_cum[i]
}
head(t)
```

```{r, eval = TRUE, message = FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
```

Some words are more frequent than others - what are the distributions of word frequencies? 

```{r, cache = TRUE, eval = FALSE}
g <- t %>% 
    filter(perc > 0.01) %>% 
    arrange(-perc) %>% 
    ggplot() + 
    geom_bar(aes(reorder(words, perc), perc), color = 'blue', stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_y_continuous(labels = scales::percent) +
    coord_flip()
g
```
How many unique words do you need in a frequency sorted dictionary to cover 50% of all word instances in the language? 
```{r, cache=TRUE, eval = FALSE}
dim(t)
t[which(t$perc_cum > .5)[1], ]
```
In this dictionary, we need 141 words over 52997 (0.26%)

90%? 
```{r, cache=TRUE, eval = FALSE}
t[which(t$perc_cum > .9)[1], ]
```
To cover 90% we need 7818 words (14.75%)

How do you evaluate how many of the words come from foreign languages? 

We can look up each word in the dataset in a English dictionary and those which are not found are from foreign languages.

Can you think of a way to increase the coverage?

We can increase the coverage adding more words or using a smaller number of words.

### n-grams

What are the frequencies of 2-grams and 3-grams in the dataset? 

The following code can calculate the frequencies of the 2-gram and 3-gram but it's 
really time consuming because it's not optimized. So we are going to use the library related with text mining and NLP instead of this code.

```{r, cache=TRUE, eval = FALSE}
# split the words line by line
splits <- strsplit(lines[inds], " ")
```

Make the n-grams by hand
```{r, cache=FALSE, eval=FALSE}
two_grams <- data.frame()
three_grams <- data.frame()
for (i in (1:length(splits))) {
    if (length(splits[[i]]) - 1 > 0) {
        word_inds <- match(splits[[i]], t$words)
        flags <- integer(length(splits[[i]]))
        flags[1] <- -1
        two_grams <- rbind(
            two_grams, 
            data.frame(
                flags, word_inds,
                fix.empty.names = FALSE
            )
        )
        if (length(splits[[i]]) - 2 > 0) {
            flags2 <- flags
            flags2[2] <- -1
            three_grams <- rbind(
                three_grams, 
                data.frame(
                    flags2, flags, word_inds,
                    fix.empty.names = FALSE
                )
            )
        }    
    }    
}
names(two_grams) <- c("w1", "w2")
names(three_grams) <- c("w1", "w2", "w3")
two_grams[two_grams$w1 == 0, "w1"] <-
    two_grams[two_grams$w1[2:(length(two_grams$w1))] == 0, "w2"][1:sum(two_grams$w1 == 0)]
two_grams <- two_grams[two_grams$w1 > 0, ]
three_grams[three_grams$w1 == 0, "w1"] <-
    three_grams[three_grams$w1[3:(length(three_grams$w1))] == 0, "w3"][1:sum(three_grams$w1 == 0)]
three_grams[three_grams$w2 == 0, "w2"] <-
    three_grams[three_grams$w2[2:(length(three_grams$w2))] == 0, "w3"][1:sum(three_grams$w2 == 0)]
three_grams <- three_grams[three_grams$w1 > 0, ]
```
```{r, cache=TRUE, eval=FALSE}
two_grams$w1 <- as.factor(two_grams$w1)
two_grams$w2 <- as.factor(two_grams$w2)

three_grams$w1 <- as.factor(three_grams$w1)
three_grams$w2 <- as.factor(three_grams$w2)
three_grams$w3 <- as.factor(three_grams$w3)
t2 <- two_grams %>%
    group_by(w1, w2) %>%
    summarise(
        freq = n(),
        perc_freq = n()/length(two_grams$w1)
    ) %>%
    arrange(-freq)
t3 <- three_grams %>%
    group_by(w1, w2, w3) %>%
    summarise(
        freq = n(),
        perc_freq = n()/length(three_grams$w1)
    ) %>%
    arrange(-freq)
```
```{r, cache=TRUE, eval=FALSE, echo=FALSE}
t2 <- table(two_grams)
# t2 <- t2[t2 > (1/100*length(two_grams))]
t2 <- t2[sort.list(as.vector(t2), decreasing = TRUE)]
# min_l <- min(length(t2), 100)
# t2 <- t2[1:min_l]
t2 <- as.data.frame(t2)
t2$perc <- t2$Freq/(sum(t2$Freq))

t3 <- table(three_grams)
# t2 <- t2[t2 > (1/100*length(two_grams))]
t3 <- t3[sort.list(as.vector(t3), decreasing = TRUE)]
# min_l <- min(length(t2), 100)
# t2 <- t2[1:min_l]
t3 <- as.data.frame(t3)
t3$perc <- t3$Freq/(sum(t3$Freq))
```

## n-grams Using Libraries

```{r, cache=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(tm)
library(RWeka)
```

```{r, cache=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
text <- VCorpus(VectorSource(lines))
```

### 1-gram

```{r, cache=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
t1 <- TermDocumentMatrix(text, control=list(tokenize = function (x) NGramTokenizer(x, control = Weka_control(min = 1, max = 1))))

f1 <- findFreqTerms(t1, lowfreq = 100)
s1 <- rowSums(as.matrix(t1[f1,]))
n1 <- data.frame(n1=names(s1), freq = s1)
```

```{r, cache=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
n1 <- n1 %>% arrange(-freq)
n1$perc <- n1$freq/(sum(n1$freq))
n1$perc_cum <- n1$perc
for (i in 2:length(n1$perc_cum)) {
    n1$perc_cum[i] <- n1$perc_cum[i - 1] + n1$perc_cum[i]
}
```

<!-- This 1-gram has 72 words because we filtered out the words with frequency less than 1000.  -->
<!-- As we saw before, this cover less than 50% of total words. So we will change the limit if we need to cover more words. -->

```{r, cache=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
g <- n1 %>% 
    filter(perc > 0.01) %>%
    arrange(-freq) %>%
    ggplot() + 
    geom_bar(aes(reorder(n1, freq), perc), color = 'blue', stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_y_continuous(labels = scales::percent) +
    coord_flip()
g
```

### 2-gram

```{r, cache=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
t2 <- TermDocumentMatrix(text, control=list(tokenize = function (x) NGramTokenizer(x, control = Weka_control(min = 2, max = 2))))

f2 <- findFreqTerms(t2, lowfreq = 100)
s2 <- rowSums(as.matrix(t2[f2,]))
n2 <- data.frame(n2=names(s2), freq = s2)
```

```{r, cache=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
n2 <- n2 %>% arrange(-freq)
n2$perc <- n2$freq/(sum(n2$freq))
n2$perc_cum <- n2$perc
for (i in 2:length(n2$perc_cum)) {
    n2$perc_cum[i] <- n2$perc_cum[i - 1] + n2$perc_cum[i]
}
```


```{r, cache=TRUE, eval = FALSE}
g <- n2 %>% 
    filter(perc > 0.005) %>%
    ggplot() + 
    geom_bar(aes(reorder(n2,freq), perc), color = 'blue', stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_y_continuous(labels = scales::percent) +
    coord_flip()
g
```

### 3-gram

```{r, cache=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
t3 <- TermDocumentMatrix(text, control=list(tokenize = function (x) NGramTokenizer(x, control = Weka_control(min = 3, max = 3))))

f3 <- findFreqTerms(t3, lowfreq = 10)
s3 <- rowSums(as.matrix(t3[f3,]))
n3 <- data.frame(n3=names(s3), freq = s3)
```

```{r, cache=TRUE, eval=FALSE, message=FALSE, warning=FALSE}
n3 <- n3 %>% arrange(-freq)
n3$perc <- n3$freq/(sum(n3$freq))
n3$perc_cum <- n3$perc
for (i in 2:length(n3$perc_cum)) {
    n3$perc_cum[i] <- n3$perc_cum[i - 1] + n3$perc_cum[i]
}
```

```{r, cache=TRUE, eval = FALSE}
g <- n3 %>% 
    filter(perc > 0.002) %>%
    ggplot() + 
    geom_bar(aes(reorder(n3,freq), perc), color = 'blue', stat = "identity") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    scale_y_continuous(labels = scales::percent) +
    coord_flip()
g
```

## Modeling

To build a basic n-gram model we can use the `markovchain` package. 

```{r, eval = TRUE, message=FALSE, warning=FALSE}
library(markovchain)
```

We need to provide the states (words, phrases or tokens) and calculate the transition matrix. We can make this model efficient because we can provide a small amount of frequent states (top words and n-grams).

To handle unseen n-grams we can use an unknown token which represents any word not included in the other states. Furthermore, we can smooth the probabilities giving a non-zero probability even if they are unseen n-grams or words.

To evaluate the model we can check the predictions given different initial states and test if the predictions have any sense.

To estimate the probability of unobserved n-grams we can use back-off models where the history is used to make predictions. Given the probabilities of the observed (n-1)-grams the probability of unseen n-grams can be estimated doing some calculations ([backoff models](https://en.wikipedia.org/wiki/Katz%27s_back-off_model)).

### Simple n-gram model
To build a simple n-gram model we will use the 2-gram frequencies as the probabilities for every word. The unseen 2-grams are given a small probability.

States
```{r, eval = FALSE}
states1 <- as.character(n1$n1)

states2 <- Reduce(paste, as.character(n2$n2))
states2 <- strsplit(states2, " ")[[1]]
states2 <- unique(states2)

states3 <- Reduce(paste, as.character(n3$n3))
states3 <- strsplit(states3, " ")[[1]]
states3 <- unique(states3)

states <- c(states1, states2, states3)
states <- unique(states)
states <- c(states, "UNK")
states <- factor(states)
states <- sort(states)

```

Transition Matrix
```{r, eval = FALSE}
m <- data.frame(row.names = states)
for (s in states) m[[s]] <- 0

for (n in 1:length(n2$n2)) {
# for (n in 1:5) {
  n22 <- strsplit(as.character(n2$n2[n]), " ")[[1]]
  m[n22[1], n22[2]] <- n2$freq[n]
}
m["UNK", ] <- 1
m$rows_sum <- rowSums(m)

m <- m %>% arrange(-rows_sum)
m <- m %>% filter(rows_sum > 0)
rnames <- row.names(m)
m <- m %>% mutate_at(vars(-rows_sum), ~ . / rows_sum) %>% select(-rows_sum)
row.names(m) <- rnames
m <- m %>% select(row.names(.))

m <- m %>% relocate(row.names(.), .after = last_col())

m <- m + 0.001
rows_sum <- rowSums(m)
m <- m %>% mutate_all(~ ./rows_sum)
row.names(m) <- rnames
```

Build the model
```{r, eval = FALSE}
mc <- new(
    "markovchain", 
    states = names(m), 
    byrow = TRUE, 
    transitionMatrix = as.matrix(m), 
    name = "n-gram Model"
)
```

Make some predictions
```{r, eval = FALSE}
preds <- c()
for (n in names(mc)) {
    w1 <- vector(mode = 'integer', length = length(names(mc)))
    w1[match(n, names(mc))] <- 1
    w2 <- w1 * (mc)
    # w3 <- w1 * (mc ^ 2)
    # preds <- rbind(preds, paste(names(mc)[which.max(w1)], names(mc)[which.max(w2)], names(mc)[which.max(w3)]))
    preds <- rbind(preds, paste(names(mc)[which.max(w1)], names(mc)[which.max(w2)]))
}
preds
```

## Prediction Model

### n-gram

```{r, cache=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
library(tm)
library(RWeka)
```

```{r, cache=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
text <- VCorpus(VectorSource(lines))
```


```{r, cache=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
t1 <- TermDocumentMatrix(text, control=list(tokenize = function (x) NGramTokenizer(x, control = Weka_control(min = 1, max = 1))))

f1 <- findFreqTerms(t1, lowfreq = 100)
s1 <- rowSums(as.matrix(t1[f1,]))
n1 <- data.frame(n1=names(s1), freq = s1)
```

```{r, cache=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
n1 <- n1 %>% arrange(-freq)
n1$perc <- n1$freq/(sum(n1$freq))
n1$perc_cum <- n1$perc
for (i in 2:length(n1$perc_cum)) {
    n1$perc_cum[i] <- n1$perc_cum[i - 1] + n1$perc_cum[i]
}
```



